---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

First load the data 

```{r load, echo=TRUE}
library(tidyverse)
df <- read_csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r mean_steps, echo=TRUE}

# find the mean steps in each day
mean_steps <- df %>%
  group_by(date) %>%
  summarize(mean_steps=mean(steps))

ggplot(mean_steps, aes(mean_steps))+geom_histogram()+labs(x="Mean number of steps per day")+theme_bw()

# find the median steps in each day

med_steps <- df %>%
  group_by(date) %>%
  summarize(Median=median(steps), Mean=mean(steps)) 

print(med_steps)

mean(med_steps$Mean, na.rm=T)
median(med_steps$Median, na.rm=T)
```

## What is the average daily activity pattern?

Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?



```{r time_plot, echo=TRUE}
time <- df %>%
  na.omit() %>%
  group_by(interval) %>%
  summarise(mean=mean(steps))

ggplot(time, aes(x=interval, y=mean))+geom_line()+labs(x="Time Interval", y= "Average steps")+theme_bw()
```

## Imputing missing values


```{r df2, echo=TRUE}
df2 <- df %>%
  group_by(interval) %>%
  mutate(steps = ifelse(is.na(steps), mean(steps,na.rm = T), steps))


# find the mean steps in each day
mean_steps2 <- df2 %>%
  group_by(date) %>%
  summarize(mean_steps=mean(steps))

ggplot(mean_steps2, aes(mean_steps))+geom_histogram()+labs(x="Mean number of steps per day")+theme_bw()

# find the median steps in each day

med_steps2 <- df2 %>%
  group_by(date) %>%
  summarize(Median=median(steps), Mean=mean(steps)) 

print(med_steps2)

mean(med_steps2$Mean, na.rm=T)
median(med_steps2$Median, na.rm=T)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r weekends, echo=TRUE}

df3 <- df %>%
  mutate(day = weekdays(date)) %>%
  mutate(which = ifelse(day %in% c("Saurday", "Sunday"), "Weekend", "Weekday")) %>%
  na.omit() %>%
  group_by(which,interval) %>%
  summarise(mean=mean(steps))

ggplot(df3, aes(x=interval, y=mean))+geom_line()+facet_wrap(~which, ncol=1)+theme_bw()+labs(x="Time interval", y= "mean steps")
```